### 2025-08-08 cognitive-bias.qmd
20250808 - regression to the mean

```{r}
#| label: fig-regressionToMean
#| fig-cap: "Regression to the Mean."
#| fig-alt: "Regression to the Mean."
#| code-fold: true

mydata <- data.frame(
  week = 1:7,
  fantasyPoints = c(25,7,12,5,11,15,3)
)

ggplot(
  data = mydata,
  mapping = aes(
    x = week,
    y = fantasyPoints
  )
) + 
  geom_point(
    size = 4
  ) +
  geom_line(
    linewidth = 1
  ) +
  geom_abline(
    intercept = 10,
    slope = 0,
    linetype = "dashed"
  ) +
  scale_x_continuous(
    breaks = 1:7
  ) +
  labs(
    y = "Fantasy Points",
    title = "Regression to the Mean"
  ) +
  annotate(
    "segment",
    x = 6.5,
    xend = 7,
    y = 17,
    yend = 10,
    color = "blue",
    linewidth = 1.5,
    alpha = 0.6,
    arrow = arrow()) +
  annotate(
    "text",
    x = 5.5,
    y = 18,
    label = "Average ability",
    hjust = 0, # left-justify
    size = 5) +
  theme_classic(base_size = 30) +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5))
```

>>>>

```{r}
#| label: fig-regressionToMean
#| fig-cap: "Regression to the Mean."
#| fig-alt: "Regression to the Mean."
#| fig-width: 10
#| fig-height: 6
#| fig-align: center
#| code-fold: true

mydata <- data.frame(
  week = 1:7,
  fantasyPoints = c(33,7,12,5,11,15,3)
)

ggplot(
  data = mydata,
  mapping = aes(
    x = week,
    y = fantasyPoints
  )
) + 
  geom_point(
    size = 4
  ) +
  geom_line(
    linewidth = 1
  ) +
  geom_abline(
    aes(
      intercept = 10,
      slope = 0,
      linetype = "Average ability"
    ),
  ) +
  scale_linetype_manual(values = "dashed") +
  coord_cartesian(
    ylim = c(0, NA)) +
  scale_x_continuous(
    breaks = 1:7
  ) +
  labs(
    y = "Fantasy Points",
    #title = "Regression to the Mean",
    linetype = NULL
  ) +
  annotate(
    "segment",
    x = 2,
    xend = 1.75,
    y = 19,
    yend = 15,
    color = "blue",
    linewidth = 1.5,
    alpha = 0.6,
    arrow = arrow()) +
  annotate(
    "text",
    x = 2,
    y = 20,
    label = "Regression toward the mean",
    hjust = 0, # left-justify
    size = 5) +
  theme_classic(base_size = 30) +
  theme(
    axis.title.y = element_text(angle = 0, vjust = 0.5),
    legend.position = "inside",
    legend.position.inside = c(0.8, 1))
```
